---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
        library(dplyr)
        library(lattice)
        rawData <- read.csv(unz("activity.zip","activity.csv"))
        
        sumStepsByDate <- summarize(group_by(rawData,date),sum(steps,na.rm=TRUE))
        names(sumStepsByDate) <- c("date","stepsPerDay")
        
        meanStepsByInt <- summarize(group_by(rawData,interval),mean(steps,na.rm=TRUE))
        names(meanStepsByInt) <- c("int","stepsPerInterval")
                         
```
## What is mean total number of steps taken per day?
## Plot the histogram of total no. of steps taken each day
```{r}
        histogram(~stepsPerDay,sumStepsByDate)
```
## 
## Calculate and report mean and median of total no. of steps per day
```{r}
        rawMean <- mean(sumStepsByDate$stepsPerDay,na.rm=TRUE)
        print(rawMean)
        rawMedian <- median(sumStepsByDate$stepsPerDay,na.rm=TRUE)
        print(rawMedian)

```
## What is the average daily activity pattern?
## Time series plot of the 5-minute interval and the average number of steps taken, averaged across 
## all days (y-axis)
```{r}
        xyplot(stepsPerInterval~int,meanStepsByInt)
        maxRowId <- which.max(meanStepsByInt$stepsPerInterval)
```
### The 5 min interval with the max number of steps (avg across all days) in the dataset is
```{r} 
print(meanStepsByInt$int[maxRowId])
```

## Imputing missing values
### Total number of missing values in the data set
```{r}
        sum(is.na(rawData))
```
### Fill in all of the missing values in the dataset using mean for that 5 min interval
```{r}
        lim1 <- 0
        lim2 <- 0
        modData <- rawData
        
        for (i in 1:61) {
                lim1 <- lim2 + 1
                if (i == 1) {lim2 <- 288} else {
                        lim2 <- lim1 + 287}
                rowCounter <- 0
                for (k in lim1:lim2) {
                  rowCounter <- rowCounter+1
                  if (is.na(rawData$steps[k])) {
                        modData$steps[k] <- meanStepsByInt$stepsPerInterval[rowCounter]
                  }
                }
        
        }
        modSumStepsByDate <- summarize(group_by(modData,date),sum(steps,na.rm=TRUE))
        names(modSumStepsByDate) <- c("date","stepsPerDay")

        histogram(~stepsPerDay,modSumStepsByDate)

        modMean <- mean(modSumStepsByDate$stepsPerDay,na.rm=TRUE)
#### Change in Mean
        print (modMean-rawMean)
        if (modMean != 0) message("Mean has changed!!!")
        modMedian <- median(modSumStepsByDate$stepsPerDay,na.rm=TRUE)
#### Change in Median
        print (modMedian-rawMedian)
        if (modMedian != 0) message("Median has changed!!!")
```        
## Are there differences in activity patterns between weekdays and weekends?
```{r}        
        x <- weekdays(as.Date(modData$date))
        
        for (i in 1:61*288) {
                if ( (x[i] == "Saturday") | (x[i] == "Sunday") ) {
                        modData$dayType[i] = "Weekend"
                } else {
                        modData$dayType[i] = "Weekdays"
                }       
        }

        modMeanStepsByInt <- summarize(group_by(modData,interval,dayType),mean(steps,na.rm=TRUE))
        names(modMeanStepsByInt) <- c("int","stepsPerInterval","dayType")
        
##        xyplot(stepsPerInterval~int | dayType,data = modMeanStepsByInt, layout = c(1,2))
        
```
